---
title: "4c. Tranformation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Normalizing, Transforming, and the QQ plot.

## 1a. Load the ChickWeight dataset from the datasets library. We will be looking at the weight distribution.

```{r library}
library(tidyverse)
library(ggplot2)
library(stats)
library(BBmisc)
library(caret)

# normalize() or scale ()
# Skew:
#   mild -> transform(sqrt()), 
#   severe -> transform(log()), 
#   BoxCoxTrans() -> predict()
```

## 1b. Plot a histogram of weight. What does the shape tell you?

It is a normal distribution.
```{r weight hist, echo=FALSE}
ChickWeight_t21 <- filter(ChickWeight, Time == 21)
print(ChickWeight_t21)
ggplot(ChickWeight_t21, aes(x=weight)) + geom_histogram()
```


## 1c. Scale the data using scale(). How does the plot of the histogram change?
It tends towards a normal distribution.
```{r scaled, echo=TRUE}
weight_scaled = scale(ChickWeight_t21$weight)
ggplot(ChickWeight_t21, aes(x=weight_scaled)) + geom_histogram()
```

## 1d. Use qqnorm() and qqline() to create quantile-quantile plots of the scaled and unscaled data. Are they different? Do you understand why?

Unscaled data:
```{r unscaled plot, echo=TRUE}
qqnorm(ChickWeight_t21$weight, main = "Unscaled Data Q-Q Plot")
qqline(ChickWeight_t21$weight)
```

Scaled data:
```{r scaled plot, echo=TRUE}
ChickWeight_t21_scaled <- mutate(ChickWeight_t21, weight = scale(weight))
qqnorm(ChickWeight_t21_scaled$weight, main = "Scaled Data Q-Q Plot")
qqline(ChickWeight_t21_scaled$weight)

```

## 1e. Try using transform() to add variables that are the log and sqrt of weight. What do you see in the quantile-quantile plots of these variables?
```{r log, echo=TRUE}
ChickWeight_t21_log = transform(ChickWeight_t21, ChickWeight_t21_log = log(weight))
qqnorm(ChickWeight_t21_log$ChickWeight_t21_log, main = "Log Q-Q Plot")
qqline(ChickWeight_t21_log$ChickWeight_t21_log)
```

```{r sqrt, echo=TRUE}
ChickWeight_t21_sqrt = transform(ChickWeight_t21, ChickWeight_t21_sqrt = sqrt(weight))
qqnorm(ChickWeight_t21_sqrt$ChickWeight_t21_sqrt, main = "Sqrt Q-Q Plot")
qqline(ChickWeight_t21_sqrt$ChickWeight_t21_sqrt)
```

## 1f. Now try using a Box-Cox transform on weight. Compare the results to those from the log and sqrt transformations.
```{r boxcox, echo=TRUE}
(ChickWeight_t21_boxcox = BoxCoxTrans(ChickWeight_t21$weight, fudge = 0.05))
ChickWeight_t21$weight_boxcox = predict(ChickWeight_t21_boxcox, ChickWeight_t21$weight)
qqnorm(ChickWeight_t21$weight_boxcox, main = "Boxcox Q-Q Plot")
qqline(ChickWeight_t21$weight_boxcox)
```

# 2. Let’s do an experiment to help us understand distributions and how understanding their shape helps us inform our transformations.

## 2a. Generate 1000 random numbers from the normal distribution with mean = 0 and sd = 0.5. Make a quantile-quantile plot.

```{r a, echo=TRUE}
(a = rnorm(1000, mean = 0, sd = 0.5))
qqnorm(a)
qqline(a)
```


## 2b. Generate 1000 random numbers from the exponential distribution with lambda = 10. Make a quantile-quantile plot.

```{r b, echo=TRUE}
(b = rexp(1000, rate = 10))
qqnorm(b)
qqline(b)
```

## 2c. Take the log of the exponentially distributed numbers. Make a quantile-quantile plot.

```{r c, echo=TRUE}
c = transform(b, b_log = log(b))
qqnorm(c$b_log)
qqline(c$b_log)
```

# 3. Bringing it all together. Scrape the mortality rate data. Examine the distribution of the mortality rate by country (histograms and quantile-quantile plots should help). Use a transformation to normalize the data if necessary.